---
author: "Copyright: Optimization Team 2022"
date: "Compiled on `r format(Sys.time(), '%B %d, %Y')`"
output: 
  prettydoc::html_pretty:
    theme: "architect"
params:
  data:
    label: "Input file:"
    value: "*.csv"
    input: text
  ministry:
    label: "Ministry Name:"
    value: AF
    input: select
    choices: [AF, BCWS, EMLI, ENV, FOR, IRR, WLRS]
  path:
    label: "Directory to Search:"
    value: "\\\\\\\\SERVERNAME.idir.bcgov\\\\SF_F[$]\\\\CONTAINER\\\\SHARE\\\\ROOT_FOLDER\\\\SUB_FOLDER"
    input: text
  folder:
    label: "Folder Name:"
    value: "SFP_Folder"
    input: text
  quarter:
    label: "Quarter:"
    value: Q1
    input: select
    choices: [Q1, Q2, Q3, Q4]
  fiscal:
    label: "Fiscal Year:"
    value: FY21-22
    input: select
    choices: [FY22-23, FY23-24]
  collected:
    label: "datestamp on OCIO file:"
    value: 2022-09-29
    input: text
title: "SFP Enhanced Report on `r params$folder` for `r params$ministry`" 
subtitle: "`r params$quarter` `r params$fiscal` (data collected on `r params$collected`)"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.path='figure/')
library(data.table)
library(DT)
library(dbplyr)
library(dplyr)
library(forcats)
library(glue)
library(here)
library(htmltools)
library(knitr)
library(lubridate)
library(openxlsx)
library(scales)
library(stringr)
library(stringi)
library(tidyverse)
library(shiny)
library(zoo)

# set working directory
here::here()
```

```{r global, call-external-functions, include = FALSE}
# load custom functions from external scripts
source(here("scripts", glue('custom_sfp_functions.R')), local = knitr::knit_global())
```

```{r load data, include = FALSE}

# load csv and fill empty cells 
sfp_data <- read_csv((here("source", glue("{params$data}", show_col_types = FALSE, na.strings=c(""," ","NULL")))))

# keep only the first 9 columns
sfp_data <- sfp_data[,c(1:9)]

# rename columns
colnames(sfp_data) <- c("filename", "path", "filetype","category","sizemb","lastaccessdate","modificationdate","creationdate","share")
```

```{r, include = FALSE}
refine_search = paste0("\\b", params$path, "\\b")
```

```{r filter results, include = FALSE}
sfp_filter1 <- filter(sfp_data, grepl({params$path}, path, ignore.case = TRUE)) 

sfp_filter <- filter(sfp_filter1, grepl(refine_search, path))
```

```{r determine file types per category, include = FALSE}
sfp_categories <- sfp_filter %>% 
  group_by(category, filetype) %>% 
  select(category, filetype)

sfp_categories_filetypes <- sfp_categories %>% 
  distinct() %>% 
  arrange(category, filetype)
```

```{r reshape data, echo = FALSE}
formatted_sfp_data <- sfp_filter %>%
  mutate(sizegb = sizemb / 1000) %>% 
  mutate(sizegb = round(sizegb, 3)) %>% 
  mutate(monthly_cost = sizegb * 2.7) %>% 
  mutate(monthly_cost = round((monthly_cost), 2)) %>% 
  mutate(obj_store_cost = sizegb * 0.07) %>% 
  mutate(object_storage_cost = round((obj_store_cost), 2)) %>% 
  mutate(lastaccessdate = convert.date(lastaccessdate)) %>% 
  mutate(modificationdate = convert.date(modificationdate)) %>% 
  mutate(creationdate = convert.date(creationdate)) %>%
  select(filename, filetype, category, path, sizemb, sizegb, lastaccessdate, modificationdate, creationdate, monthly_cost, obj_store_cost, object_storage_cost)
```

```{r determine subfolder file types per category, include = FALSE}
sfp_categories <- formatted_sfp_data %>% 
  group_by(category, filetype) %>% 
  select(category, filetype)

sfp_categories_filetypes <- sfp_categories %>% 
  distinct() %>% 
  arrange(category, filetype)
```

```{r file count all, include = FALSE}
#count total number of files
count_files <- nrow(formatted_sfp_data)
count_files_formatted <- format(count_files, big.mark = ",", scientific = FALSE)
```

```{r calculate share size, include = FALSE}
calculate_sharesize <- formatted_sfp_data %>%
  group_by(filetype) %>% 
  summarize(foldersizegb = sum(sizegb))
```

```{r calculate folder size, include = FALSE}
calculate_folder_size <- formatted_sfp_data %>%
  group_by(path) %>% 
  summarize("folder size gb" = sum(sizegb), "monthly cost" = sum(monthly_cost), "object storage cost" = sum(obj_store_cost))
```

```{r calculate category size by folder, include = FALSE}
calculate_folder_category_size <- formatted_sfp_data %>%
  group_by(path, filetype, category, monthly_cost, object_storage_cost) %>% 
  summarize(foldersizegb = sum(sizegb))
```

```{r file count per folder, include = FALSE}
#count number of files per folder
folder_filetally <- formatted_sfp_data %>%
  group_by(path) %>% 
  summarise("number of files" = n())
```

```{r calculate folder category size, include = FALSE}
folder_category_size <- formatted_sfp_data %>%
  group_by(category) %>% 
  summarize("Size(GB)" = sum(sizegb), "number of files" = n(), "monthly cost" = sum(monthly_cost)) %>% 
  select(category, "number of files", "Size(GB)", "monthly cost")

folder_category_size <- folder_category_size[order(-folder_category_size$"monthly cost"),]

folder_category_size$"monthly cost" <- dollar(folder_category_size$"monthly cost")
folder_category_size$"number of files" <- format(folder_category_size$"number of files", big.mark = ",", scientific = FALSE)
```

```{r category breakdown by sub folders, include = FALSE}
subfolder_category_size <- formatted_sfp_data %>%
  group_by(category, path) %>% 
  summarize("folder size gb" = sum(sizegb), "monthly cost" = sum(monthly_cost), "object storage cost" = sum(object_storage_cost), "number of files" = n()) %>% 
  select(path, category, "number of files", "folder size gb", "monthly cost", "object storage cost")

subfolder_category_size <- subfolder_category_size[order(-subfolder_category_size$"monthly cost"),]
```

```{r create table with folder size, depth, and file tally, include = FALSE}
table_pathsize_tallyfiles <- inner_join(calculate_folder_size,folder_filetally, by = "path") %>% 
  select(path, "number of files", "folder size gb", "monthly cost", "object storage cost")

# Count the number of '/'s in each element of string
table_pathsize_tallyfiles$folderdepth <- (str_count(table_pathsize_tallyfiles$path, "\\\\"))

# count the number of '/'s in the path parameter, subtract x from beginning of path (re: escaped backslashes)
path_depth <- stri_count_fixed(params$path, "\\") - 7

# subtract path parameter count from each path name's count for a folder depth that starts at the desired search folder
table_pathsize_tallyfiles$folderdepth <- table_pathsize_tallyfiles$folderdepth - path_depth 

table_pathsize_tallyfiles <- table_pathsize_tallyfiles[order(-table_pathsize_tallyfiles$"monthly cost"),] %>% 
  select(path, folderdepth, "number of files", "folder size gb", "monthly cost", "object storage cost")
```

```{r tally folders based on folder depth, include = FALSE}
table_folderdepth_tallyfolders <- table_pathsize_tallyfiles %>%
  group_by(folderdepth) %>% 
  tally(name = "folderdepth_count")

table_folderdepth_tallyfolders$"folderdepth_count" <- format(table_folderdepth_tallyfolders$"folderdepth_count", big.mark = ",", scientific = FALSE)
```

```{r folder count, echo = FALSE}
folder_count <- unique(formatted_sfp_data$path)
folder_count <- length(folder_count)
folder_count <- format(folder_count, big.mark = ",", scientific = FALSE)
```

```{r average folder depth, echo = FALSE}
mean_depth <- as.integer(mean(table_pathsize_tallyfiles$folderdepth))
max_depth <- as.integer(max(table_pathsize_tallyfiles$folderdepth))
```

```{r size of entire folder in GB, echo = FALSE}
folder_size <- sum(table_pathsize_tallyfiles$"folder size gb")
foldersize <- format(folder_size, big.mark = ",", scientific = FALSE)
```

```{r folder cost, include = FALSE}
foldercost <- sum(folder_size) * 2.7
folder_cost <- dollar(foldercost)
objstorecost <- sum(folder_size) * 0.07
objstore_cost <- dollar(objstorecost)
annualsavings <- sum(foldercost - objstorecost) * 12
annual_savings <- dollar(annualsavings)
```

```{r find duplicate files, include = FALSE}
# find files that are duplicated in both name and size
duplicate_files <- formatted_sfp_data %>% 
  group_by(filename, sizegb) %>% 
  filter( n() > 1 )

duplicate_files_formatted <- duplicate_files %>% 
  select (filename, path, category, sizegb, monthly_cost, lastaccessdate, modificationdate, creationdate) 

duplicate_files_formatted <- duplicate_files_formatted[order(-duplicate_files_formatted$monthly_cost),]
```

```{r duplicates cost, echo = FALSE}
# calculated at $2.70 per GB
dup_new <- duplicate_files[!duplicated(duplicate_files[ , c("filename", "sizegb")]), ]

dup_size <- round(sum(dup_new$sizegb),2)
dup_size_formatted <- format(dup_size, big.mark = ",", scientific = FALSE)

dup_cost <- dollar(dup_size * 2.7)
```

```{r count the duplicate files, include = FALSE}
dup_count <- nrow(duplicate_files_formatted) # count of all the duplicated files 
dist_dup <- nrow(dup_new) # count of each original file that has a duplicate

dup_percent <- sum((dup_count - dist_dup) / count_files) * 100
dup_percent <- format(round(dup_percent, 1), nsmall = 1)

dist_dup <- format(dist_dup, big.mark = ",", scientific = FALSE)
```

```{r find files older than 5 yrs, include = FALSE}
# convert the data collected text to date, subtract 5 years
date <- as.Date({params$collected}, "%Y-%m-%d")
dt_less5 <- ymd(date) - years(5)

# files last accessed more than 5 years ago
files_accessed_5plus <- formatted_sfp_data %>% 
  filter(lastaccessdate < dt_less5) %>% 
  select(filename, path, category, sizegb, monthly_cost, lastaccessdate) 

files_accessed_5plus <- files_accessed_5plus[(order(as.Date(files_accessed_5plus$lastaccessdate))),]


# files last modified more than 5 years ago
files_modified_5plus <- formatted_sfp_data %>% 
  filter(modificationdate < dt_less5) %>% 
  select(filename, path, category, sizegb, monthly_cost, modificationdate)

files_modified_5plus <- files_modified_5plus[(order(as.Date(files_modified_5plus$modificationdate))),]
```

```{r count the old files, include = FALSE}
la_count <- nrow(files_accessed_5plus)
lm_count <- nrow(files_modified_5plus)

la_5plus_count <- format(la_count, big.mark = ",", scientific = FALSE)
lm_5plus_count  <- format(lm_count, big.mark = ",", scientific = FALSE)

la_percent <- (la_count / count_files) * 100
lm_percent <- (lm_count / count_files) * 100

la_percent <- format(round(la_percent, 1), nsmall = 1)
lm_percent <- format(round(lm_percent, 1), nsmall = 1)
```

## SFP Information

The amount of included data in the SFP folder for `r {params$folder}` is `r {foldersize}` GB. 

The monthly cost of the included data in the `r {params$folder}` folder is **`r {folder_cost}`**.

The `r {params$folder}` folder includes `r {count_files_formatted}` files over 1MB. 

There are `r {folder_count}` folders in the `r {params$folder}` folder (excluding folders under 1MB).  

- The average folder depth for the `r {params$folder}` folder is `r {mean_depth}`, and the maximum is `r {max_depth}`. 

### Duplicate Files

Definition: same name & size as another file, located in a different sub-folder within `r {params$folder}`.  

Duplicate files make up approx. `r {dup_percent}`% of included SFP content for the `r {params$folder}` folder, totalling `r {dup_size_formatted}` GB. This represents <u>`r {dist_dup}` files</u>  with at least one copy.  
  
Approx. <span style="color: red;">`r {dup_cost}` per month</span> could be saved by cleaning up duplicates in the `r {params$folder}` folder!  

### Older Files

`r {la_percent}`% of files were last accessed over 5 years ago (`r {la_5plus_count}` files).  

`r {lm_percent}`% of files were last modified over 5 years ago (`r {lm_5plus_count}` files).    

### Data Categories
  
<div align="center"><span style="color: #234075;">**`r {params$folder}` Consumption by Data Category**</span></div>
`r datatable(folder_category_size, rownames = FALSE, colnames = c('Category' = 1, 'Number of Files' = 2, 'Size(GB)' = 3, 'Cost per Month' = 4), options=list(columnDefs = list(list(className = 'dt-left', targets="_all"))))` 
[List of category file types](https://apps.nrs.gov.bc.ca/int/confluence/download/attachments/73305882/Enhanced_Reporting_File_Categories_File_Extensions.pdf?version=1&modificationDate=1655238221000&api=v2) provided by the OCIO.

---

## About our Data Sources

*   The source data for this report is collected from the OCIO and <span style="color: red;">does not include any files under 1MB</span>. This is done by design to keep the reports manageable, as including files under 1 MB would increase the report size tenfold.

*   The OCIO data is available quarterly for analysis (Jan, Apr, Jul, Oct).

---

## SFP Reduction Goals - Records Management and Migration to Address Rising Storage Costs

- [Fiscal Year 2024-2025 Chart of NRM SFP Consumption](http://nrmccp-metabase.apps.silver.devops.gov.bc.ca/public/dashboard/4e59fe00-e957-4336-9106-e5aa9a645f78) 
- [Fiscal Year 2023-2024 Chart of NRM SFP Consumption](http://nrmccp-metabase.apps.silver.devops.gov.bc.ca/public/dashboard/709e5e4e-fd84-4653-a77d-2074b2d44110) 

- Interested in reducing SFP consumption by migrating data to Object Storage? The Optimization Team has more information on [Confluence](https://apps.nrs.gov.bc.ca/int/confluence/display/OPTIMIZE/NRM+Object+Storage+Service) or you can place a request through the [Jira Service Desk](https://apps.nrs.gov.bc.ca/int/jira/servicedesk/customer/portal/1/create/701) after reading our [Service Level Agreement](https://apps.nrs.gov.bc.ca/int/confluence/display/OPTIMIZE/Object+Storage+Forms?preview=/65671335/133730673/SLA_Revision_0.6.pdf). <br>

- [Contact the Optimization Team](mailto:NRIDS.Optimize@gov.bc.ca)

- Do you have questions about appropriate Records Management for your data? Visit [Government Records Services](https://www2.gov.bc.ca/gov/content/governments/services-for-government/information-management-technology/records-management) or [Contact the GRS Team](mailto:grs@gov.bc.ca)

```{r dynamic output name, include = FALSE}
# create output file names based on parameters
output_excel = paste0("SFP_Enhanced_Report_", params$ministry, "_", params$folder, "_", params$quarter, "_", params$fiscal, ".xlsx")
```

```{r write output to Excel file, include = FALSE}
# create workbook
excel <- createWorkbook(output_excel)

# create sheet names
firstSheet = "Folder Details"
secondSheet = "Duplicate Files"
thirdSheet = "Last Accessed 5+ Years"
fourthSheet = "Last Modified 5+ Years"
fifthSheet = "Category Details"
sixthSheet = "Category Extensions"

# add worksheets to workbook
sheet.names(firstSheet)
sheet.names(secondSheet)
sheet.names(thirdSheet)
sheet.names(fourthSheet)
sheet.names(fifthSheet)
sheet.names(sixthSheet)

# assign data tables to worksheets, apply filter across all sheets
dt.worksheets(1,table_pathsize_tallyfiles) 
dt.worksheets(2,duplicate_files_formatted)
dt.worksheets(3,files_accessed_5plus)
dt.worksheets(4,files_modified_5plus)
dt.worksheets(5,subfolder_category_size) 
dt.worksheets(6,sfp_categories_filetypes) 

# freeze top row of all sheets
freeze.panes(1)
freeze.panes(2)
freeze.panes(3)
freeze.panes(4)
freeze.panes(5)
freeze.panes(6)

# set custom column widths for all sheets
setColWidths(excel, sheet = 1, cols = c(1, 2:6), widths = c(75, 20, 20, 20, 20,20))
setColWidths(excel, sheet = 2, cols = c(1, 2, 3:8), widths = c(50, 75, 20, 20, 20, 20, 20, 20))
setColWidths(excel, sheet = 3, cols = c(1, 2, 3:8), widths = c(50, 75, 20, 20, 20, 20, 20, 20))
setColWidths(excel, sheet = 4, cols = c(1, 2, 3:8), widths = c(50, 75, 20, 20, 20, 20, 20, 20))
setColWidths(excel, sheet = 5, c(1, 2:6), widths = c(75, 20, 20, 20, 20, 20))
setColWidths(excel, sheet = 6, c(1:2), widths = c(20, 20))

# set currency format on column
sty1 = createStyle(numFmt="$0.00")
addStyle(excel, sheet = 1, sty1, rows=2:(nrow(table_pathsize_tallyfiles)+1), cols=5)
addStyle(excel, sheet = 1, sty1, rows=2:(nrow(table_pathsize_tallyfiles)+1), cols=6)
addStyle(excel, sheet = 2, sty1, rows=2:(nrow(duplicate_files_formatted)+1), cols=5)
addStyle(excel, sheet = 3, sty1, rows=2:(nrow(files_accessed_5plus)+1), cols=5)
addStyle(excel, sheet = 4, sty1, rows=2:(nrow(files_modified_5plus)+1), cols=5)
addStyle(excel, sheet = 5, sty1, rows=2:(nrow(subfolder_category_size)+1), cols=5)
addStyle(excel, sheet = 5, sty1, rows=2:(nrow(subfolder_category_size)+1), cols=6)
addStyle(excel, sheet = 6, sty1, rows=2:(nrow(sfp_categories_filetypes)+1), cols=5)

# save the workbook to file
saveWorkbook(excel, file = here("output", output_excel), overwrite = TRUE)
```