---
author: "Copyright: Optimization Team 2026"
date: "Compiled on `r format(Sys.time(), '%B %d, %Y')`"
output: 
    pdf_document:
      keep_tex: false
      df_print: kable
      highlight: breezedark
      latex_engine: xelatex
      includes:
          in_header: header.tex
urlcolor: blue
linkcolor: cyan
params:
  data:
    label: "Input file:"
    value: "*.csv"
    input: text
  ministry:
    label: "Ministry Name:"
    value: AF
    input: select
    choices: [AF, ECS, EMLI, ENV, FOR, IRR, MCM, WLRS]
  path:
    label: "Directory to Search:"
    value: "r quotation mark (Raw Data Path) quotation mark"
    input: text
  folder:
    label: "Folder Name:"
    value: "SFP_Folder"
    input: text
  collected:
    label: "datestamp on OCIO file:"
    value: 2026-03-31
    input: text
title: |
  SFP Enhanced Folder Report on `r params$folder` (`r params$ministry`)
subtitle: "Data collected on `r params$collected`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.width = 7, fig.height = 5, fig.align = "center", dev = "cairo_pdf", fig.pos = "H", fig.path = 'figure/', out.extra = "")
library(data.table)
library(DT)
library(dbplyr)
library(dplyr)
library(forcats)
library(glue)
library(here)
library(knitr)
library(lubridate)
library(magrittr)
library(openxlsx)
library(purrr)
library(scales)
library(stringr)
library(stringi)
library(tidyverse)
library(shiny)
library(zoo)
library(kableExtra)

# set working directory
here::here()
```

```{r global, call-external-functions, include = FALSE}
# load custom functions from external scripts
source(here("scripts", glue('custom_sfp_functions.R')), local = knitr::knit_global())
```

```{r load data, include = FALSE}

# load csv and fill empty cells 
sfp_data <- read_csv((here("source", glue("{params$data}", show_col_types = FALSE, na.strings=c(""," ","NULL")))))

# keep only the first 9 columns
sfp_data <- sfp_data[,c(1:9)]

# rename columns
colnames(sfp_data) <- c("filename", "path", "filetype","category","sizemb","lastaccessdate","modificationdate","creationdate","share")

# normalize the paths
sfp_data <- sfp_data %>%
  mutate(path = remap_substrings(path))
```

```{r, include = FALSE}
search_path <- remap_substrings(params$path)
refine_search <- paste0("\\b", search_path, "\\b")
```

```{r filter results, include = FALSE}

sfp_filter <- sfp_data %>%
  filter(stringi::stri_detect_fixed(path, search_path, case_insensitive = TRUE))

message("Filtered rows: ", nrow(sfp_filter))

```

```{r determine file types per category, include = FALSE}
sfp_categories <- sfp_filter %>% 
  group_by(category, filetype) %>% 
  select(category, filetype)

sfp_categories_filetypes <- sfp_categories %>% 
  distinct() %>% 
  arrange(category, filetype)
```

```{r reshape data, echo = FALSE}
formatted_sfp_data <- sfp_filter %>%
  mutate(sizegb = sizemb / 1000) %>% 
  mutate(sizegb = round(sizegb, 3)) %>% 
  mutate(monthly_cost = sizegb * 2.7) %>% 
  mutate(monthly_cost = round((monthly_cost), 2)) %>% 
  mutate(obj_store_cost = sizegb * 0.07) %>% 
  mutate(object_storage_cost = round((obj_store_cost), 2)) %>% 
  mutate(lastaccessdate = convert.date(lastaccessdate)) %>% 
  mutate(modificationdate = convert.date(modificationdate)) %>% 
  mutate(creationdate = convert.date(creationdate)) %>%
  select(filename, filetype, category, path, sizemb, sizegb, lastaccessdate, modificationdate, creationdate, monthly_cost, obj_store_cost, object_storage_cost)
```

```{r determine subfolder file types per category, include = FALSE}
sfp_categories <- formatted_sfp_data %>% 
  group_by(category, filetype) %>% 
  select(category, filetype)

sfp_categories_filetypes <- sfp_categories %>% 
  distinct() %>% 
  arrange(category, filetype)
```

```{r file count all, include = FALSE}
#count total number of files
count_files <- nrow(formatted_sfp_data)
count_files_formatted <- format(count_files, big.mark = ",", scientific = FALSE)
```

```{r calculate share size, include = FALSE}
calculate_sharesize <- formatted_sfp_data %>%
  group_by(filetype) %>% 
  summarize(foldersizegb = sum(sizegb))
```

```{r calculate folder size, include = FALSE}
calculate_folder_size <- formatted_sfp_data %>%
  group_by(path) %>% 
  summarize("folder size gb" = sum(sizegb), "monthly cost" = sum(monthly_cost), "object storage cost" = sum(obj_store_cost))
```

```{r calculate category size by folder, include = FALSE}
calculate_folder_category_size <- formatted_sfp_data %>%
  group_by(path, filetype, category, monthly_cost, object_storage_cost) %>% 
  summarize(foldersizegb = sum(sizegb))
```

```{r file count per folder, include = FALSE}
#count number of files per folder
folder_filetally <- formatted_sfp_data %>%
  group_by(path) %>% 
  summarise("number of files" = n())
```

```{r calculate folder category size, include = FALSE}
folder_category_size <- formatted_sfp_data %>%
  group_by(category) %>% 
  summarize("Size(GB)" = sum(sizegb), "number of files" = n(), "monthly cost" = sum(monthly_cost)) %>% 
  select(category, "number of files", "Size(GB)", "monthly cost")

folder_category_size <- folder_category_size[order(-folder_category_size$"monthly cost"),]

folder_category_size$"monthly cost" <- dollar(folder_category_size$"monthly cost")
folder_category_size$"number of files" <- format(folder_category_size$"number of files", big.mark = ",", scientific = FALSE)
folder_category_size$"Size(GB)" <- format(folder_category_size$"Size(GB)", digits = 3, big.mark = ",", scientific = FALSE)

folder_category_size <- folder_category_size %>% 
  rename(
    Category = category,
    "File Count" = "number of files",
    "Monthly Cost" = "monthly cost"
    )

folder_category_size <- kbl(folder_category_size, booktabs = T) %>% kable_styling(latex_options = c("striped", "hold_position"))
```

```{r create table with folder size, depth, and file tally, include = FALSE}
table_pathsize_tallyfiles <- inner_join(calculate_folder_size,folder_filetally, by = "path") %>% 
  select(path, "number of files", "folder size gb", "monthly cost", "object storage cost")

table_pathsize_tallyfiles$folderdepth <-
  stringi::stri_count_fixed(table_pathsize_tallyfiles$path, "\\") -
  stringi::stri_count_fixed(search_path, "\\")

table_pathsize_tallyfiles <- table_pathsize_tallyfiles[order(-table_pathsize_tallyfiles$"monthly cost"),] %>% 
  select(path, folderdepth, "number of files", "folder size gb", "monthly cost", "object storage cost")
```

```{r tally folders based on folder depth, include = FALSE}
table_folderdepth_tallyfolders <- table_pathsize_tallyfiles %>%
  group_by(folderdepth) %>% 
  tally(name = "folderdepth_count")

table_folderdepth_tallyfolders$"folderdepth_count" <- format(table_folderdepth_tallyfolders$"folderdepth_count", big.mark = ",", scientific = FALSE)
```

```{r folder count, echo = FALSE}
folder_count <- unique(formatted_sfp_data$path)
folder_count <- length(folder_count)
folder_count <- format(folder_count, big.mark = ",", scientific = FALSE)
```

```{r average folder depth, echo = FALSE}
mean_depth <- as.integer(mean(table_pathsize_tallyfiles$folderdepth))
max_depth <- as.integer(max(table_pathsize_tallyfiles$folderdepth))
```

```{r size of entire folder in GB, echo = FALSE}
folder_size <- sum(table_pathsize_tallyfiles$"folder size gb")
foldersize <- format(folder_size, digits = 3, big.mark = ",", scientific = FALSE)
```

```{r folder cost, include = FALSE}
foldercost <- sum(folder_size) * 2.7
folder_cost <- dollar(foldercost)
objstorecost <- sum(folder_size) * 0.07
objstore_cost <- dollar(objstorecost)
annualsavings <- sum(foldercost - objstorecost) * 12
annual_savings <- dollar(annualsavings)
```

```{r find duplicate files, include = FALSE}
# find files that are duplicated in both name and size
duplicate_files <- formatted_sfp_data %>% 
  group_by(filename, sizegb) %>% 
  filter( n() > 1 )

duplicate_files_formatted <- duplicate_files %>% 
  select (filename, path, category, sizegb, monthly_cost, lastaccessdate, modificationdate, creationdate) 

duplicate_files_formatted <- duplicate_files_formatted[order(-duplicate_files_formatted$monthly_cost),]
```

```{r duplicates cost, echo = FALSE}
# calculated at $2.70 per GB
dup_new <- duplicate_files[!duplicated(duplicate_files[ , c("filename", "sizegb")]), ]

dup_size <- round(sum(dup_new$sizegb),2)
dup_size_formatted <- format(dup_size, digits = 3, big.mark = ",", scientific = FALSE)

dup_cost <- dollar(dup_size * 2.7)
```

```{r count the duplicate files, include = FALSE}
dup_count <- nrow(duplicate_files_formatted) # count of all the duplicated files 
dist_dup <- nrow(dup_new) # count of each original file that has a duplicate

dup_percent <- if (count_files == 0) {
  0
} else {
  ((dup_count - dist_dup) / count_files) * 100
}

dup_percent <- format(round(dup_percent, 1), nsmall = 1)

dist_dup <- format(dist_dup, big.mark = ",", scientific = FALSE)
```

```{r find files older than a certain time, include = FALSE}
# convert the data collected text to date, subtract 5 years
date <- as.Date({params$collected}, "%Y-%m-%d")
dt_threshold <- file_age_threshold(
  collected_date = params$collected,
  ministry       = params$ministry
)

# files last accessed more than a certain time ago
files_accessed_old <- formatted_sfp_data %>%
  filter(lastaccessdate < dt_threshold)

files_accessed_old <- files_accessed_old[(order(as.Date(files_accessed_old$lastaccessdate))),] %>% 
  select(filename, path, category, sizegb, monthly_cost, lastaccessdate)

la_old_gb_num <- round(sum(files_accessed_old$sizegb),3)
la_old_gb <- format(la_old_gb_num, big.mark = ",", scientific = FALSE)
la_old_cost <- dollar(la_old_gb_num * 2.7)
```

```{r count the aged-out files, include = FALSE}
la_count <- nrow(files_accessed_old)

la_old_count <- format(la_count, big.mark = ",", scientific = FALSE)

la_percent <- if (count_files == 0) {
  0
} else {
  (la_count / count_files) * 100
}

la_percent <- format(round(la_percent, 1), nsmall = 1)
```

\section{SFP Information}
The `r {params$folder}` folder on SFP is \textbf{`r {foldersize}` GB}, costing the ministry **`r {folder_cost}`** per month. 

`r {params$folder}` contains `r {count_files_formatted}` files and `r {folder_count}` folders (excluding files & folders under 1MB).

\section{Duplicate Files}
\textbf{\textit{Definition:}} \textit{same name and size as another file, located within a sub-folder.}

Duplicate files over 1MB make up **`r {dup_percent}`%** of the `r {params$folder}` folder, totalling \textbf{`r {dup_size_formatted}` GB}. This represents \textbf{`r {dist_dup}` files} with at least one copy.  
  
\raggedright Approximately \fontspec[Color=Red]{Verdana Bold}`r {dup_cost}` per month \fontspec[Color=Black]{Verdana}could be saved by cleaning up duplicates in the `r {params$folder}` folder.

\section{Older Files}
**`r {la_percent}`%** of files over 1MB were last accessed a minimum of **`r file_age_label(params$ministry)`** ago \textit{(`r {la_old_count}` files)}. This represents \textbf{`r {la_old_gb}` GB} of data at a monthly cost of \fontspec[Color=Red]{Verdana Bold}`r {la_old_cost}`\fontspec[Color=Black]{Verdana}.
\clearpage

\begin{center}
\section{Data Categories}
\textbf{`r {params$folder}` SFP Consumption by Data Category}
`r {folder_category_size}` 
\href{https://apps.nrs.gov.bc.ca/int/confluence/download/attachments/73305882/Enhanced_Reporting_File_Categories_File_Extensions.pdf?version=1&modificationDate=1655238221000&api=v2}{List of category file types} provided by the OCIO.
\end{center}
\clearpage

\section{About our Data Sources}
The source data for this report is collected quarterly from the OCIO and \textcolor{red}{does not include any files under 1MB}. This is done by design to keep the reports manageable, as including files under 1 MB would greatly increase the report.

\section{SFP Reduction Goals}
#### Records management and data migration to address rising storage costs

- [Fiscal Year 2025-2026 Chart of NRM SFP Consumption](http://nrmccp-metabase.apps.silver.devops.gov.bc.ca/public/dashboard/5e12786a-2692-41de-b083-5dacc8492417) 

- \raggedright To learn more about reducing SFP consumption by migrating data to Object Storage, visit the Optimization team's [Confluence page](https://apps.nrs.gov.bc.ca/int/confluence/display/OPTIMIZE/NRM+Object+Storage+Service). You can also place a request through the [Jira Service Desk](https://apps.nrs.gov.bc.ca/int/jira/servicedesk/customer/portal/1/create/701) - please read the [Service Level Agreement](https://apps.nrs.gov.bc.ca/int/confluence/display/OPTIMIZE/Object+Storage+Forms?preview=/65671335/133730673/SLA_Revision_0.6.pdf) **before** submitting a request. 

- [Contact the Optimization Team](mailto:NRIDS.Optimize@gov.bc.ca)

`r records_management_text(params$ministry)`

```{r dynamic output name, include = FALSE}
# create output file names based on parameters
output_excel = paste0("SFP_Enhanced_Report_", params$ministry, "_", params$folder, "_", params$collected, ".xlsx")
```

```{r write output to Excel file, include = FALSE}
# create workbook
excel <- createWorkbook(output_excel)

# create sheet names
firstSheet = "Folder Details"
secondSheet = "Duplicate Files"
thirdSheet = "Older Files"
fourthSheet = "Category Extensions"

# add worksheets to workbook
sheet.names(firstSheet)
sheet.names(secondSheet)
sheet.names(thirdSheet)
sheet.names(fourthSheet)

# assign data tables to worksheets, apply filter across all sheets
dt.worksheets(1, table_pathsize_tallyfiles)
dt.worksheets(2, duplicate_files_formatted) 
dt.worksheets(3, files_accessed_old) 
dt.worksheets(4 ,sfp_categories_filetypes)

# freeze top row of all sheets
freeze.panes(1)
freeze.panes(2)
freeze.panes(3)
freeze.panes(4)

# set custom column widths for all sheets
setColWidths(excel, sheet = 1, cols = c(1, 2:6), widths = c(75, 20, 20, 20, 20, 20))
setColWidths(excel, sheet = 2, cols = c(1, 2, 3:8), widths = c(50, 75, 20, 20, 20, 20, 20, 20))
setColWidths(excel, sheet = 3, cols = c(1, 2, 3:8), widths = c(50, 75, 20, 20, 20, 20, 20, 20))
setColWidths(excel, sheet = 4, cols = c(1:2), widths = c(20, 20))

# set currency format on column
sty1 = createStyle(numFmt="$0.00")
addStyle(excel, sheet = 1, sty1, rows=2:(nrow(table_pathsize_tallyfiles)+1), cols=5)
addStyle(excel, sheet = 1, sty1, rows=2:(nrow(table_pathsize_tallyfiles)+1), cols=6)
addStyle(excel, sheet = 2, sty1, rows=2:(nrow(duplicate_files_formatted)+1), cols=5)
addStyle(excel, sheet = 3, sty1, rows=2:(nrow(files_accessed_old)+1), cols=5)

# save the workbook to file
saveWorkbook(excel, file = here("output", output_excel), overwrite = TRUE)
```